---
title: "Exploring U.S. Supreme Court Decisions"
author: "Kejing Li"
output: github_document
---

## Get the data

```{r, echo = TRUE}
# load useful packages
library(tidyverse)
library(lubridate)
# load data
scdbv_mod <- read_csv("data/SCDB_2019_01_justiceCentered_Citation.csv")
scdbv_leg <- read_csv("data/SCDB_Legacy_05_justiceCentered_Citation.csv")
scdbv_mod
scdbv_leg

theme_set(theme_minimal())
```

## Combine the datasets

```{r}
#convert docket and adminActionState into numerical 
scdbv_leg <- 
  scdbv_leg %>%
  mutate(docket = as.numeric(docket),
       adminActionState = as.numeric(adminActionState)) 

#combine the datasets
scdbv <- bind_rows(scdbv_leg, scdbv_mod)
```

## Recode variables as you find necessary

```{r}
#select variables for analysis purpose 
scdbv_select <- scdbv%>% 
  select(caseIssuesId, term ,justice, justiceName, decisionDirection, majVotes, minVotes, majority, chief, dateDecision, decisionType)

scdbv_select
```

## What percentage of cases in each term are decided by a one-vote margin (i.e. 5-4, 4-3, etc.)

```{r}
scdbv_select %>%
#calculate the margin
  drop_na(majVotes, minVotes) %>%
  mutate(margin = majVotes - minVotes) %>%
  group_by(term) %>% 
#calculate the percent
  summarize(percent = sum(margin == 1)/n())%>%
  ggplot(aes(term, percent))+
  geom_line()+
#change the y-axis into percent format
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  labs(title = "Percent of U.S. Supreme Court cases decided by 1-vote margin",
         x = "Term", 
         y = "Percent of total cases decided",
         caption = "Source: The Supreme Court Database")  

```

__Answer:__ As seen above, the percent of total cases decided by 1-vote margin peaked at 50% around 1800, and stayed relatively constant afterwards until 1925, where an increase trend was shown. 


## In each term he served on the Court, in what percentage of cases was Justice Antonin Scalia in the majority?

```{r}
scdbv_select %>%
#Justice Antonin Scalia is coded as AScalia in the dataset
  filter(justiceName == "AScalia") %>%
  group_by(term) %>%
#majority is coded 2 in the dataset 
  summarize(percent = sum(majority == 2, na.rm = TRUE)/n()) %>%
  ggplot(aes(term, percent))+
  geom_line()+
#change the y-axis into percent format and set its range from 0 to 1.
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                     limits = c(0, 1))+
  labs(title = "Percent of the time Scalia was in the majority",
         x = "Term", 
         y = "Percentage of total decisions",
         caption = "Source: The Supreme Court Database") 

```

__Answer:__ In terms he served on the Court, Justice Antonin Scalia was in the majority in over 70% cases.

## Create a graph similar to above that adds a second component which compares the percentage for all cases versus non-unanimous cases (i.e. there was at least one dissenting vote)

```{r}
scdbv_select %>%
  filter(justiceName == "AScalia") %>%
  group_by(term) %>%
  summarize(percent = sum(majority == 2, na.rm = TRUE)/n(),
#calculate the percent in the non-unanimous cases
#minVotes is coded as the number of votes in dissent
            percent_nu = sum(majority == 2 & minVotes >= 1, na.rm = TRUE)/sum(minVotes >= 1)) %>%
  ggplot()+
#use color to distinguish two lines 
  geom_line(aes(term, percent, color = "All decisions"))+
  geom_line(aes(term, percent_nu, color = "Non-unanimous decisions"))+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
labs(title = "Percent of the time Scalia was in the majority",
         x = "Term", 
         y = "Percentage of total decisions",
         caption = "Source: The Supreme Court Database",
         color = "") +
  theme(legend.position = "bottom")

```

__Answer:__ The graph suggests that Scalia was more likely to stay with majority when there was no dissenting votings. 

## In each term, what percentage of cases were decided in the conservative direction?

```{r}
scdbv_select %>%
  group_by(term) %>%
#conservative decision is coded 1 in the dataset
  summarize(percent = sum(decisionDirection == 1, na.rm = TRUE)/n()) %>%
  ggplot(aes(term, percent))+
  geom_line()+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 0.7))+
  labs(title = "U.S. Supreme Court",
         subtitle = "Percent of cases decided in a conservative direction",
         x = "Term", 
         y = "Percent of total decisions",
         caption = "Source: The Supreme Court Database")

```

__Answer:__ In the given terms, percentage of cases decided in the conservative direction fluctrated between 20% and 60%.


## In each term, how many of the term's published decisions (decided after oral arguments) were announced in a given month?

```{r}
#set month levels
#reverse because `coord_flip` is deployed in the following code
month_levels <- rev(c("Oct", "Nov", "Dec","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep"))

scdbv_select %>%
#decided after oral arguments is coded 1, 6, 7 in the dataset 
  filter(decisionType == c(1, 6, 7)) %>%
#change the data type of dataDecision into date
  mutate(dateDecision = mdy(dateDecision)) %>%
#extract month component from date Decision
  mutate(month = month(dateDecision)) %>%
#group by term-month
  group_by(term, month) %>%
#count by each distinct case id
  summarize(numbers = n_distinct(caseIssuesId)) %>%
#set levels and label them
  ggplot(aes(factor(month, 
                    levels = rev(c(10, 11, 12, 1, 2, 3, 4, 5, 6, 7, 8, 9)),
                    labels = month_levels), numbers))+
  geom_boxplot()+
  coord_flip()+
  labs(title = "U.S. Supreme Court",
         subtitle = "Number of decisions announced post-oral arguments per month, by term",
         x = "", 
         y = "Number of decisions announced in a term-month",
         caption = "Source: The Supreme Court Database")
```

__Answer:__ As shown above, the third quarter saw least number of decision announced post-oral arguments, while the second quarter, especially June, was where most published decisions were annonuced through out the year. 

## Session info

```{r, echo = TRUE}
devtools::session_info()
```


